---
title: "Video Games Analysis"
output: html_notebook
---

You have been hired by Donkey Games, a games company located in Dundee, Scotland. They are planning their next release and want to use data to decide what kind of game to make. They have a dataset that contains historical information on the top selling games of all time.

Donkey want to understand what types of games sell a lot of copies. In particular, they are looking for analysis that helps them decide what kind of games their company should make in the future
. 
```{r}

library(dplyr)

library(ggplot2)
```

```{r}
# check the working directory
getwd()

df <- read.csv("game_sales.csv", header = TRUE,sep = ",")
df
```
```{r}
#Checking for null values
colSums(is.na(df))
```
```{r}
summary(df)
```
##Data MAnipilation
We will try to find insights based on the data
###Genre Analysis
```{r}
table(df$year_of_release)
```

```{r}
#get the value counts for the genre
table(df$genre)

```
```{r}
genre_counts <- table(df$genre)
genre_data <- as.data.frame(genre_counts)
colnames(genre_data) <- c("Genre", "Count")
# Create the pie chart
pie_chart <- ggplot(genre_data, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity") +
  coord_polar("y") +  # Convert to a pie chart
  theme_void()  # Remove unnecessary elements
pie_chart

```


```{r}
df2 <- df # Duplicate the data
df2$genre_mean <- ave(df$sales,df$genre) #Get the mean sales of the genres and save in a new column
df2
```
```{r}
library(ggplot2)
ggplot(df2, aes(x = genre, y = genre_mean)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.7) +
  labs(title = "Mean Sales by Genre",
       x = NULL,
       y = "Average Sales") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())+
  facet_wrap(~ genre, scales = "free_x", labeller = label_both)
```
```{r}
ggplot(
  head(df, 10),
  aes(
    x=sales,
    y=reorder(name, sales),
    fill=year_of_release
  )
) +
  geom_bar(stat="identity") +
  labs(
    y="Name",
    x=" Sales"
  )
```
##publisher analysis

```{r}
genre_counts <- table(df$publisher)
genre_data <- as.data.frame(genre_counts)
colnames(genre_data) <- c("Genre", "Count")
# Create the pie chart
pie_chart <- ggplot(genre_data, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity") +
  coord_polar("y") +  # Convert to a pie chart
  theme_void()  # Remove unnecessary elements
pie_chart
```

```{r}
publisher_sales <- aggregate(
  sales~publisher+year_of_release,
  df,
  sum
)

```

```{r}
publisher_sales_clean <- aggregate(
  sales~publisher,
  df,
  sum
)
publisher_sales_clean <- publisher_sales_clean[
  order(publisher_sales_clean$sales, decreasing=T),
]
ggplot(
  publisher_sales,
  aes(
    x=sales,
    y=reorder(publisher, sales),
    fill=year_of_release
  )
) +
  geom_bar(stat="identity") +
  scale_fill_continuous(low="red", high="blue") +
  scale_y_discrete(limits=head(publisher_sales_clean, 10)$publisher) +
  labs(
    y="Platform",
    x="Sales"
  )

```
##Year Analysis
```{r}
# Scatter plot of Global Sales vs. Critic Score
ggplot(df, aes(x = year_of_release, y = sales)) +
  geom_point() +
  labs(title = " Sales vs. year",
       x = "year",
       y = "Sales ")
```



```{r}
ggplot(
  df,
  aes(
    x=year_of_release,
  )
) +
  geom_histogram(
    binwidth=1,
    fill="green"
  ) +
  theme_minimal()

```
##platform analysis
```{r}
bar_chart <- ggplot(df, aes(x = platform, fill = "Red")) +
  geom_bar() +
  scale_fill_manual(values = "red") +  # Set fill color to blue
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
bar_chart
```
```{r}
# Scatter plot of Global Sales vs. Critic Score
ggplot(df, aes(x = critic_score, y = sales)) +
  geom_point() +
  labs(title = " Sales vs. Critic Score",
       x = "Critic Score",
       y = "Sales ")
```
```{r}
# Scatter plot of Sales vs.user Score
ggplot(df, aes(x = user_score, y = sales)) +
  geom_point() +
  labs(title = "Sales vs. user Score",
       x = "User Score",
       y = "Sales (in millions)")
```
```{r}
# Extract correlation coefficients for Global Sales and User Score
cor_sales_user <- correlation_matrix["sales", "user_score"]

# Print correlation coefficients
cat("Correlation between Sales and critic score:", cor_sales_user, "\n")
```
```{r}
# Extract correlation coefficients for Sales and Critic Score
cor_sales_critic <- correlation_matrix["sales", "critic_score"]

cat("Correlation between Sales and User Score:", cor_sales_critic, "\n")
```



```{r}
# Calculate correlation matrix
correlation_matrix <- cor(df[c("sales", "critic_score", "user_score")])

# Print correlation matrix
print(correlation_matrix)
```



```{r}

# Create a heatmap to visualize the correlation
ggplot(data = melt(correlation_matrix), aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), vjust = 1) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Correlation") +
  theme_minimal() +
  labs(title = "Correlation Heatmap",
       x = "Variables",
       y = "Variables")


```
```{r}
#Regression analysis
# Fit a simple linear regression model
regression_model <- lm(sales ~ critic_score, data = df)

specific_critic_score <- 80

new_data <- data.frame(critic_score = specific_critic_score)
predicted_sales <- predict(regression_model, newdata = new_data)
cat("Predicted Sales for Critic Score", specific_critic_score, ":", predicted_sales, "\n")
```

